Exploratory_Analysis_Josh_Tobey

### 2025-12-08 Exploratory_Analysis_Josh_Tobey.Rmd
Working on GAM Fit.

coagual

>>>>

```{r}
ggplot(exploratory_df) +
  geom_point(aes(x = mean_pH,
                 y = mean_Temp))
```
```{r}
RWAR_dosing_lag_incdec <- read_csv("RWAR_dosing_lag_incdec.csv")
head(RWAR_dosing_lag_incdec)

```
```{r}
RWAR_dosing_lag_incdec <- RWAR_dosing_lag_incdec %>%
  select(-c(FeCl_change, Polymer_change, Timestamp))
head(RWAR_dosing_lag_incdec)
```

```{r}
library(mgcv)

lm_spec <- linear_reg()

gam_formula <- Ferric.chloride..mg.L ~ s(Cationic.Polimer..mg.L) + s(mean_Temp) + s(mean_pH) + s(mean_Conductivity) + s(mean_Turbidity) + s(Temp_lag1) + s(Temp_lag2) + s(pH_lag1) + s(pH_lag2) + s(Cond_lag1) + s(Cond_lag2) + s(Turb_lag1) + s(Turb_lag2)

gam_fit <- gam(gam_formula, data = RWAR_dosing_lag_incdec)
gam_rec <- recipe(Ferric.chloride..mg.L ~ ., data = RWAR_dosing_lag_incdec) %>%
  step_ns(Cationic.Polimer..mg.L, deg_free = 10) %>%
  step_ns(mean_Temp, deg_free = 10) %>%
  step_ns(mean_pH, deg_free = 10) %>%
  step_ns(mean_Conductivity, deg_free = 10) %>%
  step_ns(mean_Turbidity, deg_free = 10) %>%
  step_ns(Temp_lag1, deg_free = 10) %>%
  step_ns(Temp_lag2, deg_free = 10) %>%
  step_ns(pH_lag1, deg_free = 10) %>%
  step_ns(pH_lag2, deg_free = 10) %>%
  step_ns(Cond_lag1, deg_free = 10) %>%
  step_ns(Cond_lag2, deg_free = 10) %>%
  step_ns(Turb_lag1, deg_free = 10) %>%
  step_ns(Turb_lag2, deg_free = 10)

gam_wf <- workflow() |>
  add_model(lm_spec) |>
  add_recipe(gam_rec)

gam_fit <- gam_wf |>
  fit(data = RWAR_dosing_lag_incdec)

gam_fit |> tidy()
gam_fit |> glance()

dat_grid <- tibble(expand.grid(Cationic.Polimer..mg.L = seq(min(RWAR_dosing_lag_incdec$Cationic.Polimer..mg.L), 
                               max(RWAR_dosing_lag_incdec$Cationic.Polimer..mg.L)),
                               mean_Temp = seq(min(RWAR_dosing_lag_incdec$mean_Temp), max(RWAR_dosing_lag_incdec$mean_Temp)),
                               mean_pH = seq(min(RWAR_dosing_lag_incdec$mean_pH), max(RWAR_dosing_lag_incdec$mean_pH)),
                               mean_Conductivity = seq(min(RWAR_dosing_lag_incdec$mean_Conductivity), max(RWAR_dosing_lag_incdec$mean_Conductivity)),
                               mean_Turbidity = seq(min(RWAR_dosing_lag_incdec$mean_Turbidity), max(RWAR_dosing_lag_incdec$mean_Turbidity)),
                               Temp_lag1 = seq(min(RWAR_dosing_lag_incdec$Temp_lag1), max(RWAR_dosing_lag_incdec$Temp_lag1)),
                               Temp_lag2 = seq(min(RWAR_dosing_lag_incdec$Temp_lag2), max(RWAR_dosing_lag_incdec$Temp_lag2)),
                               pH_lag1 = seq(min(RWAR_dosing_lag_incdec$pH_lag1), max(RWAR_dosing_lag_incdec$pH_lag1)),
                               pH_lag2 = seq(min(RWAR_dosing_lag_incdec$pH_lag2), max(RWAR_dosing_lag_incdec$pH_lag2)),
                               Cond_lag1 = seq(min(RWAR_dosing_lag_incdec$Cond_lag1), max(RWAR_dosing_lag_incdec$Cond_lag1)),
                               Cond_lag2 = seq(min(RWAR_dosing_lag_incdec$Cond_lag2), max(RWAR_dosing_lag_incdec$Cond_lag2)),
                               Turb_lag1 = seq(min(RWAR_dosing_lag_incdec$Turb_lag1), max(RWAR_dosing_lag_incdec$Turb_lag1)),
                               Turb_lag2 = seq(min(RWAR_dosing_lag_incdec$Turb_lag2), max(RWAR_dosing_lag_incdec$Turb_lag2))
                               ))
bind_cols(
  augment(gam_fit, new_data = dat_grid),
  predict(gam_fit, new_data = dat_grid, type = "conf_int")
) -> dose_pred_gam

for(pred in c("age", "year", "education")) {
  pred_dat <- dose_pred_gam[, c(pred, ".pred", ".pred_lower", ".pred_upper")] |>
    group_by_at(pred) |>
    summarise(.pred = mean(.pred), .pred_lower = mean(.pred_lower), .pred_upper = mean(.pred_upper))
  dose_dat <- RWAR_dosing_lag_incdec[, c(pred, "Ferric.chloride..mg.L")]

  if(is.numeric(dose_dat |> pull(pred))) {
    p <- dose_dat |>
      ggplot() +
      geom_point(aes_string(pred, "wage"), alpha = .3) +
      geom_line(aes_string(pred, ".pred"), data = pred_dat, colour = "darkgreen") +
      geom_line(aes_string(pred, ".pred_lower"), data = pred_dat, colour = "darkgreen", lty = 2) +
      geom_line(aes_string(pred, ".pred_upper"), data = pred_dat, colour = "darkgreen", lty = 2)
  } else {
    p <- dose_dat |>
      ggplot() +
      geom_jitter(aes_string(pred, "wage"), alpha = .3) +
      geom_point(aes_string(pred, ".pred"), data = pred_dat, colour = "darkgreen") +
      geom_errorbar(aes_string(pred, ymin = ".pred_lower", ymax = ".pred_upper"), data = pred_dat, colour = "darkgreen")
  }
  print(p)
}
